Plant-pollintor networks worldwide are composed by the same specific building

### 2021-06-16 Draft/Draft.Rmd
First full description of funtional goups with their respecting names added

Motifs are abstract representations often decoupled from species ecology. Hence, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear if species ecological and life history traits strongly determines its role within the network of interactions [@coux2016]. For example, bees are more specialized foragers [Jose, aquí por ejemplo, pueden decir que abejas pequeñas y generalistas, pueden comportarse muy diferente que grandes y especialistas... si tienes una lista de pollintors, con su familia, pasamela, y le hecho un ojo](), usually alble to open complex zygomorphic flowers which are not accessible to fly pollinators. How this different behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], they are often overlooked in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

@@

cite complexheatmap https://doi.org/10.1093/bioinformatics/btw313

@@

Next, we calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the ‘bmotif’ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. To fit the LMMs models in our analyses, we used the ‘lmer’ package [@bates2015].

@@

The hierarchical cluster analysis divided the dataset with 1506 species and 15 traits into five different clusters with different and overlapping characteristics (Figure S1). The subset of plant species used in this study (N = 503) were distributed evenly across these different larger five groups (see Figure S1). The first cluster "short-lived selfers"

The first and largest cluster, Cluster A, was constituted by 710 species. The species of this cluster had the shortest plant heights (Median = 0.5m, Mean = 0.84m), were mostly herbs (89.58%) with short lived life span (68.23%) and the lowest number of flowers per plant (Median = 30, Mean = 197.15); the flowers were mostly hermaphrodite (97.40%) with actinomorphic symmetry (99.48%), the morphology was variant but the predominant shapes were ‘open’ (51.56%) and ‘capitulum’ (23.44%); most of the species were self compatible (96.35%) and had the higher levels of autonomous selfing of all clusters from low-medium selfing (47.92%) to high selfing (52.08%) with nectar production for 83.85% of the species. Cluster B, had 588 species and was the the second in plant height (Median = 0.8m, Mean = 1.98m), number of flowers per plant (Median = 75; Mean = 1559) and flowers per inflorescence (Median = 15; Mean = 58), it had diverse life forms (herbs 52.89%, shrubs 34.52% and trees 12.59%) and the species were mostly perennial (97.96%); the flowers were the second largest (flower length Median = 10mm, Mean = 10.08mm; flower width Median = 5.10mm, Mean = 16.88mm) and it had mostly hermaphroditic breeding system (95.41%), actinomorphic symmetry (99.15%) and a predominant flower morphology of ‘open’ (50.17%) and ‘tube’ shapes (20.07); the majority of species were self compatible (70.24%) and partially self compatible (20.24%) and had low selfing levels (88.95%) with nectar production for 78.40%. Cluster C was constituted by 323 species, had similar plant heights to cluster B (Median = 0.75, Mean = 1.85) with also diverse life forms (61.61% herbs, 28.79% shrubs, 9.60% trees) with mostly perennial species (95.36%) with the largest flower size (flower length Median =9.10mm , Mean =13.95mm; flower width Median =15mm, Mean =20.18mm) and number of ovules per flower of all clusters (Median = 30, Mean  = 527.87); flowers were mostly hermaphroditic (97.83%) with actinomorphic flowers (78.95%) and predominant ‘open’ (28.48%) ‘capitulum’ (27.55%) and ‘tube’ (26.01%) flower morphology; this was the cluster with greater proportion of self incompatible species (76.47%) and it had negligible autonomous selfing level (Median = 0,  Mean = 0.30) with the highest proportion of nectar secreting species (92.26%). Cluster D had 164 species, they were the tallest plants (Median =2.88m, Mean =5.19m) and had the largest proportion of shrubs (41.46%) and trees (31.71%) of all clusters, the species were mainly perennial (92.07%) and differentiated from the rest of the clusters by having principally dioecious (63.41%) and monoecious (35.98%) breeding system; this cluster had the largest number of flower per plant (Median = 1000, Mean = 6803) and per inflorescence (Median = 30 , Mean = 258) but the smaller flower size of all clusters (flower length Median =2.71, Mean =6.53; flower width Median = 3.62; Mean = 8.23) with ‘open’ (53.66%) and ‘brush’ (21.34%) morphology as the most representative flower shapes of this group; in addition, this cluster had the shortest styles (Median = 1mm; Mean = 3.05mm) and lower number of ovules (Median = 4.50, Mean = 110.47) along with cluster A; thus, the flowers were unisexual (98.78%), actinomorphic (95.12%), had no selfing (98.17%) and had the lowest proportion of nectar secreting species (69.51%) of all clusters. Finally, 239 species integrated Cluster E which had the second shortest species of all clusters (Median = 0.55m; Mean = 1.23m), the species were mainly herbs (83.26%) and perennial (70.29%) but also had a representative proportion of short lived species (29.71%); this cluster was the only one with mainly zygomorphic flowers (97.07%) with tubular (55.65%) and papilionaceous (24.69%) flower shapes, the species had mostly hermaphroditic breeding system (96.65%) and the flowers had the longest styles of all clusters (Median = 5mm; Mean = 8.74mm); species were mostly self compatible (88.28%) and nectar secreting (88.28%) with low autonomous selfing levels (61.92%) but also had species with high (18.41%) medium (15.06%) and lack of autonomous selfing (4.60%).

Figure S1

@@

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Short-lived selfers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Short-lived selfers",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))

>>>>

Motifs are abstract representations often decoupled from species ecology. Hence, linking the structural properties of the meso-scale with the species ecology can help the understanding of ecological processes. For instance, different motifs can have different ecological meanings [@simmons2019] and the position within a motif can determine the species functional role [@stouffer2012; @baker2015]. However, it is unclear if species ecological and life history traits strongly determines its role within the network of interactions [@coux2016]. For example, large pollinators can forage larger distances [@greenleaf2007], deposit greater pollen quantities [@foldesi2021] and handle complex zygomorphic flowers in comparison with small pollinators that are restricted to lower floral complexity [@gong2009]. How this different pollinator behaviors translate into their interaction topology is unknown. Similarly, recent empirical findings indicate that the meso-scale is the best descriptor of plant reproductive success [@allen2021], but little is known on how plants reproductive strategies shape their position within the network of interactions. Although some studies have evaluated plant reproductive strategies in plant-pollinator networks [@tur2013; @lazaro2020], they are often overlooked in a community context [@devaux2014] and rarely incorporated into plant-pollinator network studies. Hence, exploring how the main plant reproductive strategies integrate with the emergent motif framework can shed light on key aspects of ecosystem functioning.

@@

@@

Next, we calculated which functional groups were over or under-represented in different motif positions by comparing position frequencies of empirical networks with those of their corresponding simulated counterparts. We extracted the position frequencies of all motifs from three to five nodes for each network and species by using the ‘bmotif’ package [@bmotifsimmons2019]. To estimate the position frequencies of each functional group in a given network, we added the frequencies of those species that belong to the group, and then, we normalised the resulting frequencies by dividing the position measure for each group by the total number of times that a group appears in any position within the same motif size class. Then, we calculated the percentile of the observed position frequencies for each group and network, just like we did motif frequencies. To outline the general patterns of position frequencies across networks and functional groups, we fit a LMM per motif position, where the response variable was the observed position percentile per network. We used the functional group identifier as an explanatory variable and the study identifiers as a random intercept. By adding the functional group estimates to the model intercept, we assessed the average motif frequency, after controlling the variation at the study level. To fit the LMMs models in our analyses, we used the ‘lmer’ package [@bates2015]. Finally, we visualized with the help of the package ComplexHeatmap version 2.6.2 [@gu2016] over- and under- representation of plant and pollinator functional groups on the different motif positions.

@@

The hierarchical cluster analysis divided the dataset with 1506 species and 15 traits into five different clusters with different and overlapping characteristics (Figure S1 and Figure S2). The subset of plant species used in this study (N = 503) were distributed evenly across these different larger five groups (see subset of species labels within cluster in Figure S1). The first cluster was dominated by herbs with hermaphrodite flowers with high levels of autonomous selfing and we we refer to this group as "selfing herbs". The second cluster was characterized by small perennial species with a mixed of life forms (trees shrubs and herbs) with outcrossing hermaphroditic flowers and we refer to this group of species as "small outcrossing perennials". The third cluster was dominated by also perennial species with a mixed of life forms too and had large self-incompatible hermaphroditic flowers with high number of ovules and we named this group "self-incompatible perennials with large flowers". The fourth cluster had the tallest species, highest proportion of shrub and tree life forms, dioecious and monoecious breeding system with small flowers but the highest number of flowers per plant and inflorescence and we refer to this set of species as "tall plants with small unisexual flowers". Finally, the last cluster of species was dominated by small perennial and shot-lived herbs with long self-compatible zygomorphic flowers unable to self-pollinate and we refer to this group as "short-lived outcrossers with long zygomorphic flowers".

@@

long_plants_1$Functional_groups[long_plants_1$Functional_groups==1] <- "Selfing herbs" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==2] <- "Small outcrossing\nperennials" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==3] <- "Self-incomp. perennials\nwith large flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==4] <- "Tall plants with small\nunisexual flowers" 
long_plants_1$Functional_groups[long_plants_1$Functional_groups==5] <- "Short-lived outcrossers with\nlong zygomorphic flowers"

long_plants_1$Functional_groups <- factor(long_plants_1$Functional_groups, levels=c("Selfing herbs",
"Small outcrossing\nperennials" ,"Self-incomp. perennials\nwith large flowers",
"Tall plants with small\nunisexual flowers" ,"Short-lived outcrossers with\nlong zygomorphic flowers" ))
